# Lowess Practical

## 1. Generate simulated data

```{r}
#| echo: true

set.seed(1)
x <- 1:100
y <- sin(x/10) + rnorm(100,0,0.2)
```

## 2. Implement the LOWESS Algorithm

```{r}
#| echo: true

customLowess <- function(x,y,f){ 
  k <- ceiling(f*100)
  y_hat <- numeric(100) #wghts <- c() #ngb <- x[1:k]
  for (i in 1:100){ 
    # find the nearest k neighbors 
    distances <- abs(x-x[i]) 
    sortedIndices <- sort.list(distances) 
    nearest_indices <- sortedIndices[1:k]

    x_ngb <- x[nearest_indices]
    y_ngb <- y[nearest_indices]

    dmax <- max(distances[nearest_indices])
    wghts <- (1- (distances[nearest_indices]/dmax)^3)^3
    W <- diag(wghts)

    X <- cbind(1,x_ngb)
    beta <- solve(t(X)%*%W%*%X)%*%t(X)%*%W%*%y_ngb
    y_hat[i] <- c(1,x[i])%*%beta
  } 
  return(y_hat) 
}
```

### 2.1 Comparing values from different functions

We now need to check if the custom function yields the same results as the in-built R lowess() function. For this we render the following code:

```{r}
#| echo: true

customLowess(x,y,0.1)
lowess(x,y, f = 0.1, iter = 0)
```

## 3. Plots

We then plot the original data and the smoothed data for comparison:

```{r}
#| echo: true

smooth_y <- customLowess(x,y,0.1)
plot(y~x)
plot(smooth_y ~ x)

```

From the plots, we can see that the lowess function produces less scattered values compared to the original data points.
